---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {r2eng}

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

The goal of the {r2eng} (as in 'R to English') package is to take an R expression and 'translate' it to an English sentence. Make R speakable!

The project inspired by [Amelia McNamara](https://twitter.com/AmeliaMN)'s useR! 2020 talk ([YouTube](https://www.youtube.com/watch?v=ckW9sSdIVAc), [slides](https://www.amelia.mn/SpeakingR/#1)).

This project is a work in progress and highly opinionated.

## Installation

You can install the development version of {r2eng} from GitHub with:

```{r install, eval=FALSE}
remotes::install_github("matt-dray/r2eng")
```

This package depends on {lintr} and {purrr}.

## Examples

There's currently one function in the package: `r2eng()`.

Pass it an R expression like this:

```{r example}
library(r2eng)
r2eng::r2eng(variable <- 1, speak = TRUE)
```

Set speak to `TRUE` for a system call that will read the English sentence out loud.

```{r example2}
obj <- r2eng::r2eng(hello <- c(TRUE, FALSE, 'banana' %in% c('apple', 'orange')), speak = FALSE)
obj
```

An `r2eng` has two methods: `speak` to speak the English sentence.

```{r example3}
speak(obj)
```

`evaluate` to evaluate the expression.

```{r example3eval}
evaluate(obj)
hello
```

More complicated examples.

```{r example4}
r2eng::r2eng(mtcars %>% select(mpg > 22) %>% mutate(gear4 = gear == 4), speak = FALSE)
```

```{r example5}
r2eng::r2eng(ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + geom_smooth(), speak = FALSE)
```

It can understand the meaning of `=` when used for assignment versus specifying arguments. But for some expressions (e.g. those using `=` for variable assignment), you must use `r2eng_from_string`. ^[r2eng uses [non-standard evaluation](http://adv-r.had.co.nz/Computing-on-the-language.html). Feeding an expression such as `x = c(1, 2, 3)` would confuse `r2eng` that you want to pass an argument `c(1, 2, 3)` to the parameter `x`. This package has been opinionated: It highlights using = for variable assignment is not a good idea.]

```{r example6}
r2eng::r2eng_from_string("x = c(1, 2, 3)", speak = FALSE)
```

```{r example7}
r2eng::r2eng(plot(x = c(1, 2, 3), y = c(5, 6, 7)), speak = FALSE)
```

## r2eng versus r2eng\_from\_string

The Raison d'être for `r2eng\_from\_string` is to handle some exceptional cases, e.g. expression using `=` for variable assignment above. It is also useful when you pipe an expression.

```{r pipe}
library(magrittr)
"non_english <- c('ceci n est pas une pipe', 'Ich bin ein Berliner', '我其實唔識講廣東話')" %>% r2eng_from_string(speak = TRUE)
```

## Work in progress (WIP)

There is much to do. Most R expressions won't currently work with the `r2eng()` function.

- [ ] Expand the dictionary
- [X] Split out parentheses for evaluation
- [ ] Ensure multi-line translation
- [X] Smart check of expression structure (e.g. '=' will be used as gets if used for assignment, but will be 'is' elsewhere)
- [ ] Allow for variant opinions on translations
- [ ] Account for dialects (dollar, formula, tidyverse, etc, notation)
- [ ] Test!
- [ ] Add vignettes

## Code of Conduct

I welcome contributions.

Please note that the {r2eng} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
