---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {r2eng}

<!-- badges: start -->
<!-- badges: end -->

The goal of the {r2eng} package is to take an R expression and 'translate' it to an English sentence. 

The project inspired by [Amelia McNamara](https://twitter.com/AmeliaMN)'s useR! 2020 talk ([YouTube](https://www.youtube.com/watch?v=ckW9sSdIVAc), [slides](https://www.amelia.mn/SpeakingR/#1)).

The project is a work in progress.

## Installation

You can install the released version of r2eng from GitHub with:

```{r install, eval=FALSE}
remotes::install_github("r2eng")
```

There are no dependencies.

## Example

There's currently one function in the package: `r2eng()` (as in 'R to English').

Pass it an R expression like this:

```{r example}
r2eng::r2eng("variable <- 1", speak = FALSE)
```

Set speak to `TRUE` for a system call that will read the English sentence out loud.

## WIP

There is much to do. Most R expressions won't currently work with the `r2eng()` function.

- [ ] Expand dictionary
- [ ] Split out parentheses
- [ ] Smart check of expression structure, so '=' will be used as gets if used for assignment, but will be 'is' elsewhere 
- [ ] Allow for variant opinions on translations
- [ ] Vignette

## Code of Conduct

I welcome contributions.

Please note that the {r2eng} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
