---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {r2eng}

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

The goal of the {r2eng} (as in 'R to English') package is to take an R expression and 'translate' it to an English sentence. Make R speakable!

The project inspired by [Amelia McNamara](https://twitter.com/AmeliaMN)'s useR! 2020 talk ([YouTube](https://www.youtube.com/watch?v=ckW9sSdIVAc), [slides](https://www.amelia.mn/SpeakingR/#1)).

This project is a work in progress and highly opinionated.

## Installation

You can install the development version of {r2eng} from GitHub with:

```{r install, eval=FALSE}
remotes::install_github("matt-dray/r2eng")
```

This package depends on lintr and purrr.

## Example

There's currently one function in the package: `r2eng()`.

Pass it an R expression like this:

Set speak to `TRUE` for a system call that will read the English sentence out loud.

```{r example}
library(r2eng)
r2eng::r2eng("variable <- 1", speak = TRUE)
```


```{r example2}
obj <- r2eng::r2eng("hello <- c(TRUE, FALSE, 'banana' %in% c('apple', 'orange'))", speak = FALSE)
obj
```

```{r example3}
speak(obj)
```

```{r example4}
r2eng::r2eng("mtcars %>% select(mpg > 22) %>% mutate(gear4 = gear == 4)")
```

```{r example5}
r2eng::r2eng("ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + geom_smooth()")
```

It can understand the meaning of =

```{r example6}
r2eng::r2eng("x = c(1, 2, 3)")
```

```{r example7}
r2eng::r2eng("plot(x = c(1, 2, 3), y = c(5, 6, 7))")
```

## Work in progress (WIP)

There is much to do. Most R expressions won't currently work with the `r2eng()` function.

- [ ] Expand the dictionary
- [X] Split out parentheses for evaluation
- [ ] Ensure multi-line translation
- [X] Smart check of expression structure (e.g. '=' will be used as gets if used for assignment, but will be 'is' elsewhere)
- [ ] Allow for variant opinions on translations
- [ ] Account for dialects (dollar, formula, tidyverse, etc, notation)
- [ ] Test!
- [ ] Add vignettes

## Code of Conduct

I welcome contributions.

Please note that the {r2eng} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
