---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {r2eng}

ɑː ˈtuː /eng/

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

Make R speakable!

The goal of {r2eng} (as in 'R to English') is to take an R expression and 'translate' it to an English sentence. 

The package is intended to:

* improve communication between teachers and learners
* make R discussions in English more accessible to non-English speakers
* provide an extra audio evaluation tool for users who are blind or have low vision
* be of interest to any R user that's curious about how R expressions might be vocalised

The project was inspired by [Amelia McNamara](https://twitter.com/AmeliaMN)'s useR! 2020 talk called 'Speaking R' ([YouTube](https://www.youtube.com/watch?v=ckW9sSdIVAc), [slides](https://www.amelia.mn/SpeakingR/#1)).

This project is a work in progress and highly opinionated. Contributions are welcome, but please see the [Code of Conduct](#conduct).

## Installation

You can install the development version of {r2eng} from GitHub with:

```{r install, eval=FALSE}
remotes::install_github("matt-dray/r2eng")
```

This package depends on {lintr}, {purrr} and {rlang}.

## Examples

The main function in the package is `r2eng()`. It uses [non-standard evaluation](http://adv-r.had.co.nz/Computing-on-the-language.html), so you pass it a bare R expression like this:

```{r example}
r2eng::r2eng(variable <- 1, speak = TRUE)
```

Set speak to `TRUE` for a system call that will read the English sentence out loud.

A more complex example:

```{r example2}
obj <- r2eng::r2eng(
  hello <- c(TRUE, FALSE, 'banana' %in% c('apple', 'orange')),
  speak = FALSE
)

obj
```

### Methods

An `r2eng` object has two methods: `speak` and `evaluate`.

Use `speak` to speak the English sentence.

```{r example3}
r2eng::speak(obj)
```

Use `evaluate` to evaluate the expression.

```{r example3eval}
r2eng::evaluate(obj)
hello
```

From your r2eng object you can access the original R expression (`r_expression`), English translation (`eng_expression`) and quoted expression (`quoted_expression`). You can also see the parse tree output via {lintr}:

```{r map}
head(obj$translation_map)
```

### Further examples

Here's an example using the pipe (`%>%`) and two types of 'equals':

```{r example4}
library(magrittr)
r2eng::r2eng(
  mtcars %>% filter(mpg > 22) %>% mutate(gear4 = gear == 4),
  speak = FALSE
)
```

This example uses the 'plus' constructor from {ggplot2}:

```{r example5}
r2eng::r2eng(
  ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + geom_smooth(),
  speak = FALSE
)
```

This example shows what happens when you pass vectors:

```{r example6}
r2eng::r2eng(plot(x = c(1, 2, 3), y = c(5, 6, 7)), speak = FALSE)
```

### Passing a string

The `r2eng()` function understands the meaning of `=` when used for assignment versus specifying arguments, but feeding an expression such as `x = c(1, 2, 3)` would confuse `r2eng()` that you want to pass an argument `c(1, 2, 3)` to the parameter `x`. 

This is because `r2eng()` uses [non-standard evaluation](http://adv-r.had.co.nz/Computing-on-the-language.html). 

In such cases, you must use `r2eng_from_string()` instead:

```{r example7}
r2eng::r2eng_from_string("x = c(1, 2, 3)", speak = FALSE)
```

Another exceptional case for `r2eng_from_string()` is when piping and expression:

```{r pipe}
"non_english <- c('ceci n est pas une pipe', 'Ich bin ein Berliner', '我其實唔識講廣東話')" %>% 
  r2eng::r2eng_from_string(speak = TRUE)
```

## RStudio addin

Installing this package also installs an [RStudio addin](https://rstudio.github.io/rstudioaddins/).

Select an R expression in the editor and then under 'Addins', go to 'Vocalise R expression' under 'R2ENG'. The selected text will be spoken by your computer. 

You could bind this addin to a keyboard shortcut in RStudio by going to 'Tools', then 'Modify Keyboard Shortcuts...'. Perhaps <kbd>Cmd</kbd> + <kbd>Shift</kbd> + <kbd>V</kbd>.

As with the rest of this package, vocalisation is only possible for macOS and with speakers.

## Work in progress (WIP)

There is much to do.

- [ ] Expand the dictionary
- [X] Split out parentheses for evaluation
- [ ] Ensure multi-line translation
- [X] Smart check of expression structure (e.g. '=' will be used as gets if used for assignment, but will be 'is' elsewhere)
- [ ] Allow for variant opinions on translations
- [ ] Account for dialects (dollar, formula, tidyverse, etc, notation)
- [ ] Test!
- [ ] Add documentation (vignettes, {pkgdown} site)

## Code of Conduct {#conduct}

Contributions are welcome from everyone.

Please note that the {r2eng} project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
